Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-09-06 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:09:06	22:37:29	(Sun 6.Sep)

\author{Jo~Inge~Buskenes, %
        Andreas~Austeng, %
        Carl-Inge~Columbo~Nilsen%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the Department
of Informatics, University of Oslo, Norway.}% <-this % stops a space

@@

Adaptive beamformers process each pixel in a different way in an attempt to surpass the conventional limit of resolution and noise suppression. This non-linear processing of the image increases its bandwidth and introduces a need for oversampling. For MVDR often a factor 10 lateral oversampling is often needed to ensure spatial shift-invariance in source amplitude~\cite{Asen2014}. To compare apples to apples, we will use the same factor for both LCA and MVDR.

@@

Idea to steer within a fraction of the rectangular window's mainlobe.

\begin{align}
d = \frac{\pi\,\theta_\text{open}\,\lambda}{M\,d\,O_s}
\end{align}

@@

Assuming that a robust MVDR computes the optimal windows in the minimum variance sense, it seems sensible to create similar ones for LCA. We study this in \Fig{mvdr_selected_windows}, where an MVDR sectorscan image of the Holmengraa wreck is presented, along with the window responses MVDR chose in the shadow, speckle and highlight regions of this image.

@@

If we look at the peak sidelobe levels of the MVDR windows they are for the most part between -13\,dB and -43\,dB. This corresponds to that of the rectangular and Hamming window, or a Kaiser window with $\beta=0$ and $\beta=2$. Choosing another value for $\beta$ will  This suggests This and symmetry suggests

Since the responses as similar to those 
All the responses in \Fig{mvdr_selected_windows} can be approximated with the Kaiser-Bessel function. By adjusting the window parameter $\beta$ we can adjust the trade-off between mainlobe width and sidelobe level. 
To generate windows with different mainlobe width and sidelobe level,

A parametric window like Kaiser-Bessel allow us to create approximately these and the responses in between.

The Kaiser-Bessel windows s with $\beta$-values of 0 and 2 allows us to approximately trade between these two boundaries. \todo{add the advantages here?}

In our studies

The Kaiser windows are easy to compute, is fairly insensitive to coefficient inaccuracies, and can create windows with shapes spanning from rectangular to Gaussian.

@@

\subsection{Kaiser parameter boundaries}

@@

\newcommand\cc[1]{\multicolumn{1}{>{\columncolor{tabBlue}}c}{\bf #1}}
\begin{table}[!b]\centering%\normalsize
\begin{tabular}[c]{l r r r@{}  r r r}\hline
\rowcolor{tabBlue} & \cc{Resolution} & \cc{S} & \cc{$\beta$} & \cc{$\varphi^*$} &  \cc{N$_\beta$} & \cc{N$_\varphi$} \\\hline
& 50\%  & 28 & [0,3] & \hspace{.15cm} 50\% & 2 & 3 \\
& 50\%  & 28 & [0,6] & \hspace{.15cm} 50\% & 2 & 3 \\
& 50\%  & 28 & [0,9] & \hspace{.15cm} 50\% & 2 & 3 \\
& 100\% & 32 &  &:30 \\
& 150\% & 32 & 1 &:4\\
\end{tabular}\\
* Steering is relative to the -3dB width of the window being applied.
\caption{Qualitative study}\label{parameter_study}
\end{table}%

@@

@@

\subsection{Summary and order of results}

In short, our findings regarding the optimal window types for LCA is as follows:

\begin{itemize}
\item \emph{Window function} should be symmetric, . Studying typical MVDR responses for shadow, speckle and highlight regions, we believe it to be insignificant
\item \emph{Window parameters}. 
\end{itemize}

% \begin{itemize}
% \item Sell it!
% \item As briefly as possible, introduce beamforming.
% \item Adaptive beamformer's potential lies in its ability to suppress interference power
% \item Why adaptive beamformers struggle in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive. Constraints must be applied in one way or another - parameters must be tuned.
% \begin{itemize}
% \item Noise and signal is correlated. Spatial averaging required. 
% \item Increases variance of speckle (not only in active systems?), spatial compunting \cite{Vignon2009} (US) or time averaging?. Subarray averaging applied here \cite{Synnevag2007} (US) \todo{Is it bad to mix ultrasound/ sonar refs?}
% \end{itemize}
% \item LCA is like DAS, only we test several windows and choose the one that yield the lowest beamformer output power. Ultrasound \cite{Synnevag2008}, active sonar \cite{Buskenes2011}\cite{Blomberg2011}\cite{Blomberg2012}\todo{who was that guy in the 80ties?}
% \item Outline: 
% \begin{itemize}
% \item Methods.
% \item See what happens to Capon's response %\cite{Synnevag2007}  when means for robustification are added.
% \item Setup the window set for LCA to match this.
% \item Notice that the LCA performs like Capon, even with a small window database.
% \item Start with Kaiser, see that other window functions also work well.
% \end{itemize}
% \end{itemize}
% \todopar{Out of the blue idea: Weigh CF with LCA? This guy did it with MVDR \cite{Wang2009}. Probably you know already... Okay, back on track.}

\section{Methods}

% \todopar{Threw some formulas in to get a feel of them}

\begin{align}
z[n] = \sumb{m=0}{M-1} w_m[n]^*x_m[n-\Delta_m] = \w\H[n]\Xd[n]
\end{align}
where
\begin{align}
\w[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]} \quad \text{and} \quad\Xd[n] = \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]}.
\end{align}

Minimum variance distortionless resonse \cite{Capon1969}
\begin{align}
\underset{\w[n]}{\argmin}\, E\{ |z[n]|^2 \} = \underset{\w[n]}{\argmin}\, \w\H[n]\R[n]\w[n], 
\end{align}
subject to
\begin{align}
\w\H[n]\a = 1,
\end{align}
where
\begin{align}
\R[n] = E\{ \x\x\H \}.
\end{align}

\section{Results}

% \begin{figure}[!t]\centering
% % \includegraphics[width=2.5in]{}
% \caption{Simulation Results}
% \label{fig_sim}
% \end{figure}

See end of document...
% \begin{figure*}[!t]
% \centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
% \label{fig_first_case}}
% \hfil
% \subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
% \label{fig_second_case}}}
% \caption{Simulation results}
% \label{fig_sim}
% \end{figure*}

\begin{itemize}
\item Something similar to the abstract.
\end{itemize}

% \ \\
% \IEEEPARstart{T}{o} form images from a modern phased array sonar system the received wavefield is usually recorded, and then postprocessed by a digital beamformer. The beamformer applies delays and weights to the sensor channels, the beamformer adjusts the arrays spatial response to focus at one pixel at a time.  such that signals emanating from regions of interest add constructively, while ensuring that noise and interference from other angles do not. 
% 
% The imaging capabilities of a modern phased array sonar system depend on physical attributes such element response and array geometry, the transmitted signal, as well as the beamforming method being used on transmission and reception. Beamforming is the concept of applying delays and weights to the sensors channels to steer the arrays response to points of interest.

% 
% 
% Outline:
% \begin{itemize}
% \item Capon's resonse when applying robustification
% \item Choice of window functions makes little difference.
% \item Steering and mainlobewidths have outer bounds.
% \item Beamspace?
% \item Chosen window plots - what may they tell us? Variance intensity values when using various windows.
% \item Assymmetric windows?
% \end{itemize}
% 
% 
% \begin{align}
% z[n] = \sumb{m=0}{M-1} w_m[n]^*x_m[n-\Delta_m] = \w\H[n]\x[n-\Delta_m]
% \end{align}
% 
% 
% \section{Methods}
% 
% Basically, we are working with a practical implementation of the Capon beamformer that computes a set of weights $\vec w$ for every single sample $n$ by solving:
% \begin{gather*}
% \vec w[n] = \frac{\hat{\mat R}\,\!^{-1}[n]\vec a}{\vec a\H\hat{\mat R}\,\!^{-1}[n]\vec a} = \frac{\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}{\vec a\H\text{\raisebox{1.9pt}{$\vec\chi$}}[n]}
% \end{gather*}%
% where
% % \newcommand\X{\text{\raisebox{2pt}{$\vec\chi$}}}
% \begin{gather*}
% \text{\raisebox{1.9pt}{$\vec\chi$}} = \hat{\mat R}\,\!^{-1}\vec a \qquad\qquad\text{is the solution to}\qquad\qquad \hat{\mat R}\text{\raisebox{1.9pt}{$\vec\chi$}} = \vec a.
% \end{gather*}

% 
% , and maximum suppression of while ensuring that the beamformer digitally  before each of the pixels are estimated one at a time. The resolution and contrast of such a system will depend on the systems spatial response, which ideally should be narrow  be very sharp in the desired direction its ability to achieve  fundamental principle of forming a sonar image is to record the received wavefield, 
% 
% image quality of a phased array sonar imaging systems depend on  the choice of weights to apply to each of the sensors are crucial. 
% 
% A modern phased array imaging system may be thought of as a spatial filter. To achieve the best possible performance, the 
% 
% resolution and contrast 
% 
% Adaptive beamformers have only recently been introduced in active sonar imaging. For a while they were considered unsuited for this purpose because the backscattered signal is largely correlated with the 
% 
% 
%

%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
% \begin{figure*}[!t]
% \centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
% \label{fig_first_case}}
% \hfil
% \subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
% \label{fig_second_case}}}
% \caption{Simulation results}
% \label{fig_sim}
% \end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% However, Computer Society journals sometimes do use bottom floats - bear
% this in mind when choosing appropriate optional arguments for the
% figure/table environments.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Conclusion}
The conclusion goes here.

We limit our study to the use of variations of the parametric Kaiser-Harris window, since we observed neglible performance difference to using this function or any of the common trigonometric ones. The Kaiser windows are easy to compute, is fairly insensitive to coefficient inaccuracies, and can create windows with shapes spanning from rectangular to Gaussian.

Using the LCA beamformer results in increased robustness and improved edge definition compared to the MV and DAS beamformers.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

@@

\bibliographystyle{IEEEtran}
\bibliography{references}
% 
% 
% \input{bio/jo_inge_buskenes}
% % \input{bio/jon_petter_aasen}
% % \input{bio/carl-inge_nilsen}
% \input{bio/andreas_austeng}

>>>>

\author{Jo~Inge~Buskenes, %
        Roy~Edgar~Hansen,%
        Andreas~Austeng%
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem All authors are with the Department
of Informatics, University of Oslo, Norway.}% <-this % stops a space

@@

The ultimate goal of adaptive beamformers is to take the information available in the wavefield and use it to either improve image resolution, noise suppression, or both. This non-linear processing increases the image bandwidth and hence introduces a to oversample it to capture all the details. For MVDR a factor 10 oversampling laterally compared to the Nyquist rate is often needed to ensure  spatially local shift-invariance in source amplitude~\cite{Asen2014}. Since LCA operates with the same optimization criterion, we will be using approximately a factor 8 lateral oversampling for both beamformers.

@@

% Idea to steer within a fraction of the rectangular window's mainlobe.

% \begin{align}
% d = \frac{\pi\,\theta_\text{open}\,\lambda}{M\,d\,O_s}
% \end{align}

@@

Assuming that a robust MVDR is the reference method we want LCA to perform similarly to, it seems sensible to create window database for LCA with responses that resembles that of MVDR. We study this in \Fig{mvdr_selected_windows}, where an MVDR sectorscan image of the Holmengraa wreck is presented, along with the window responses MVDR chose in the shadow, speckle and highlight regions of this image.

@@

If we look at the peak sidelobe levels of the MVDR windows they are for the most part between -13\,dB and -43\,dB. This corresponds to that of the rectangular and Hamming window, or a Kaiser window with $\beta=0$ and $\beta=2$, respectively. Using the Kaiser window in this case makes sense since it is optimal in the sense of having its peak power concentration near the steering angle, and because we can also create any response in between. The Kaiser windows are easy to compute, is fairly insensitive to coefficient inaccuracies, and can create windows with shapes spanning from rectangular to Gaussian.

@@

\subsection{Kaiser parameter $\beta$ and steering $\phi$}

@@

% \newcommand\cc[1]{\multicolumn{1}{>{\columncolor{tabBlue}}c}{\bf #1}}
% \begin{table}[!b]\centering%\normalsize
% \begin{tabular}[c]{l r r r@{}  r r r}\hline
% \rowcolor{tabBlue} & \cc{Resolution} & \cc{S} & \cc{$\beta$} & \cc{$\varphi^*$} &  \cc{N$_\beta$} & \cc{N$_\varphi$} \\\hline
% & 50\%  & 28 & [0,3] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 50\%  & 28 & [0,6] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 50\%  & 28 & [0,9] & \hspace{.15cm} 50\% & 2 & 3 \\
% & 100\% & 32 &  &:30 \\
% & 150\% & 32 & 1 &:4\\
% \end{tabular}\\
% * Steering is relative to the -3dB width of the window being applied.
% \caption{Qualitative study}\label{parameter_study}
% \end{table}%

@@

% We limit our study to the use of variations of the parametric Kaiser-Harris window, since we observed neglible performance difference to using this function or any of the common trigonometric ones. The Kaiser windows are easy to compute, is fairly insensitive to coefficient inaccuracies, and can create windows with shapes spanning from rectangular to Gaussian.
% 
% 
% Using the LCA beamformer results in increased robustness and improved edge definition compared to the MV and DAS beamformers.

@@

% \begin{itemize}
% \item Adaptive beamformer's potential lies in its ability to suppress interference power
% \item Why adaptive beamformers struggle in active sonar systems. Correlated noise, robustification kills the adaptive potential. Quite computationally intensive. Constraints must be applied in one way or another - parameters must be tuned.
% \begin{itemize}
% \item Noise and signal is correlated. Spatial averaging required. 
% \item Increases variance of speckle (not only in active systems?), spatial compunting \cite{Vignon2009} (US) or time averaging?. Subarray averaging applied here \cite{Synnevag2007} (US) \todo{Is it bad to mix ultrasound/ sonar refs?}
% \end{itemize}
% \item Outline: 
% \begin{itemize}
% \item Setup the window set for LCA to match this.
% \item Notice that the LCA performs like Capon, even with a small window database.
% \item Start with Kaiser, see that other window functions also work well.
% \end{itemize}
% \end{itemize}
% \todopar{Out of the blue idea: Weigh CF with LCA? This guy did it with MVDR \cite{Wang2009}. Probably you know already... Okay, back on track.}

% \IEEEPARstart{T}{o} form images from a modern phased array sonar system the received wavefield is usually recorded, and then postprocessed by a digital beamformer. The beamformer applies delays and weights to the sensor channels, the beamformer adjusts the arrays spatial response to focus at one pixel at a time.  such that signals emanating from regions of interest add constructively, while ensuring that noise and interference from other angles do not. 
% 
% The imaging capabilities of a modern phased array sonar system depend on physical attributes such element response and array geometry, the transmitted signal, as well as the beamforming method being used on transmission and reception. Beamforming is the concept of applying delays and weights to the sensors channels to steer the arrays response to points of interest.

% 
% 
% Outline:
% \begin{itemize}
% \item Choice of window functions makes little difference.
% \item Steering and mainlobewidths have outer bounds.
% \item Beamspace?
% \item Chosen window plots - what may they tell us? Variance intensity values when using various windows.
% \item Assymmetric windows?
% \end{itemize}
%

% , and maximum suppression of while ensuring that the beamformer digitally  before each of the pixels are estimated one at a time. The resolution and contrast of such a system will depend on the systems spatial response, which ideally should be narrow  be very sharp in the desired direction its ability to achieve  fundamental principle of forming a sonar image is to record the received wavefield,

@@

\bibliographystyle{IEEEtran}
\bibliography{references}
% 
% 
\input{bio/jo_inge_buskenes}
\input{bio/andreas_austeng}
